Verifying Concurrent Search Structure Templates for C Programs using VST

### 2023-04-26 concurrency/CONCUR 2023/paper.tex
fix some breakage

\paragraph*{Related Work}
\than{I think we will write related work inside introduction part}

@@

\subsection{Proofs of Templates} \than{Can someone suggest another title?}

In this section, we verify \lstinline{traverse} for both lock-coupling and give-up templates.

@@

@@

\subsection{Templates vs. Internal Reorganization}
Our original goal was to verify an existing concurrent BST implementation with the template approach, refactoring as necessary but otherwise leaving the code unchanged. However, there is a fundamental mismatch between the template approach and the standard approach to deletion in a BST. The operations we have seen thus far can be factored into a data-structure-agnostic concurrent step (\lstinline{traverse}) and a concurrency-unaware data structure step (\lstinline{insert_op} or \lstinline{lookup_op}). When removing a node from a BST, on the other hand, we usually restructure the tree with a \lstinline{pushdown} operation, %include figure?
rotating three-node sections of the tree until the node to be deleted is at a leaf. This operation is simultaneously concurrency-aware (we must lock the nodes involved to avoid race conditions) and data-structure-specific (we need to know precisely which nodes to target and how to rearrange them). This contrasts with linked-list deletion, which only involves the target node and its immediate neighbors. In practice, the existing template proofs for tree structures either do not include deletion at all, or use a very lightweight form of deletion in which nodes are marked as ``deleted'' but are never actually removed from the data structure.

We believe this represents a fundamental limitation in the template approach. Operations in the style of \lstinline{pushdown} are ubiquitous in tree structures (e.g., the rebalancing operations of AVL and red-black trees), and always weave together synchronization and restructuring in a way that cannot easily be decomposed. This is a serious impediment to the goal of verifying concurrency control separately from the data structure: even if we have a verified template and a verified single-threaded tree, we still need to verify a complex data-structure-specific concurrent function. %in fact, usually the most complicated part!
%what's the conclusion here? can we see a way through?

\section{Proof Mechanization}
We present the mechanized formal verification of the correctness of these templates, which are implemented in C and verified with the Verified Software Toolchain. We employ concurrent separation logic to analyze the behavior of the C code and incorporate Iris-style logical atomicity within VST.

>>>>

\subsection*{Related Work}
\than{I think we will write related work inside introduction part}

\than{Write more here}

@@

\subsection{Proofs of Templates} \
%than{Can someone suggest another title?}
In this section, we verify \lstinline{traverse} for both lock-coupling and give-up templates.

@@

Our primary focus is on verification the correctness of \texttt{insert} operations in binary search trees using lock-coupling and give-up templates. The \texttt{lookup} operations closely resemble the \texttt{insert} operations, sharing a significant portion of their proof. As for the \texttt{delete} operations, we have supplied proof solely for the lock-coupling template, which is discussed in Section \ref{internal_reorganize}.

The complete proof can be found online in our publicly accessible repository of machine-checked proofs (refer to Section \ref{proof_mech}). This repository also contains the other proofs featured in our monograph.

@@

\subsection{Templates vs. Internal Reorganization}
\label{internal_reorganize}
Our original goal was to verify an existing concurrent BST implementation with the template approach, refactoring as necessary but otherwise leaving the code unchanged. However, there is a fundamental mismatch between the template approach and the standard approach to deletion in a BST. The operations we have seen thus far can be factored into a data-structure-agnostic concurrent step (\lstinline{traverse}) and a concurrency-unaware data structure step (\lstinline{insert} or \lstinline{lookup}). When removing a node from a BST, on the other hand, we usually restructure the tree with a \lstinline{pushdown} operation, %include figure?
rotating three-node sections of the tree until the node to be deleted is at a leaf. This operation is simultaneously concurrency-aware (we must lock the nodes involved to avoid race conditions) and data-structure-specific (we need to know precisely which nodes to target and how to rearrange them). This contrasts with linked-list deletion, which only involves the target node and its immediate neighbors. In practice, the existing template proofs for tree structures either do not include deletion at all, or use a very lightweight form of deletion in which nodes are marked as ``deleted'' but are never actually removed from the data structure.

We believe this represents a fundamental limitation in the template approach. Operations in the style of \lstinline{pushdown_left} are ubiquitous in tree structures (e.g., the rebalancing operations of AVL and red-black trees), and always weave together synchronization and restructuring in a way that cannot easily be decomposed. This is a serious impediment to the goal of verifying concurrency control separately from the data structure: even if we have a verified template and a verified single-threaded tree, we still need to verify a complex data-structure-specific concurrent function. %in fact, usually the most complicated part!
%what's the conclusion here? can we see a way through?

\section{Proof Mechanization}
\label{proof_mech}
We present the mechanized formal verification of the correctness of these templates, which are implemented in C and verified with the Verified Software Toolchain. We employ concurrent separation logic to analyze the behavior of the C code and incorporate Iris-style logical atomicity within VST.
